---
title: "Archaeology-Exploration"
format: html
---

#The Beginning of the Exploration of the Data of Archaeologists

```{r setup, include=FALSE}
knitr::knit_engines$set(sparql =SPARQLchunks::eng_sparql) #set the package SPARQLchunks as default application for SPARQL
```
```{r, include=FALSE}
library(SPARQLchunks)
```

##General exploration of the List of Archaeologists on Wikipedia via DBpedia:

1. The number of Archaeologists in the LOA (List of Archaeologists)
```{sparql output.var="number-LOA", endpoint="https://dbpedia.org/sparql", warning="FALSE"}
PREFIX dbr: <http://dbpedia.org/resource/> #Definition of the value dbr: as a prefix!
PREFIX dbo: <http://dbpedia.org/ontology/>
select (count(*) as ?number)
where {
dbr:List_of_archaeologists dbo:wikiPageWikiLink ?s.
?s a dbo:Person
}
LIMIT 10
```
2. 15 examples of Archaeologists in the LOA
```{sparql output.var="examples-LOA", endpoint="https://dbpedia.org/sparql", warning="FALSE"}
select ?s
where {
dbr:List_of_archaeologists dbo:wikiPageWikiLink ?s.
?s a dbo:Person
}
LIMIT 15
```
3. The highest most often available attributes of the List of Archaeologists:
```{sparql output.var="attributes-LOA", endpoint="https://dbpedia.org/sparql", warning="FALSE"}
PREFIX dbr: <http://dbpedia.org/resource/>
PREFIX dbo: <http://dbpedia.org/ontology/>
SELECT ?p1 (COUNT(*) as ?eff)
WHERE { 
dbr:List_of_astronomers ?p ?o1.
?o1 a dbo:Person;
    ?p1 ?o2.
  }
GROUP BY ?p1
ORDER BY DESC(?eff)
```
In the Wikipedia List of Archaeologists a total of 'r number-LOA$number' Archaeologists is represented in the semantic web. Several attributes scraped of the Wikipedia infoboxes are available for many of the data points: What they were known for ('r attributes-LOA$p1=="http://dbpedia.org/ontology/knownFor"'), their Birth Place ('r attributes-LOA$p1=="http://dbpedia.org/ontology/birthPlace"'), awards ('r attributes-LOA$p1=="http://dbpedia.org/ontology/award"') and their academic discipline ('r attributes-LOA$p1=="http://dbpedia.org/ontology/academicDiscipline"').




