---
title: "Introduction-SPARQLchunks"
format: html
---

#Introduction to SPARQL using the SPARQLchunks package

The SPARQLchunks package ([view setup here](Setup-SPARQL.qmd)) allows the query of SPARQL endpoints from R, R Markdown and Quarto environments. Accessed data can both be shown as lists and and stored in data frames.

To ensure proper workability of the package, following content needs to be included in the R Markdown setup chunk:
knitr::knit_engines$set(sparql = SPARQLchunks::eng_sparql)
```{r setup, include=FALSE}
knitr::knit_engines$set(sparql =SPARQLchunks::eng_sparql) #set the package SPARQLchunks as default application for SPARQL
```

##Access SPARQL via a SPARQL chunk options:

To access SPARQL via the SPARQLchunks package we first need to create a sparql code block. In the first line several things can be defined:
- output.var="name": name of the data frame in which data should be stored.
- endpoint="url.com": url of the endpoint that is utilized for the query.
- the SPARQL code is usable in other SPARQL access points. 
```{sparql output.var="result_df", endpoint="https://example.com/sparql"}
PREFIX ex: <http://example.com/schema/> #Definition of the value ex: as a prefix!
SELECT ?subject ?predicate ?object
WHERE {
  ?subject ?predicate ?object.
}
LIMIT 10
```

##Access SPARQL in R code:

It is furthermore possible to embed SPARQL queries in R code using the package. This can therefore be integrated into a workflow that accesses different databases before an analysis. By defining the value of Variables with the endpoint url or the SPARQL query the command sparql2df allows us to convert a SPARQL query into an instantly usable data frame. To cross-reference this code with a different program, the content of the query variable and the endpoint have to be extracted manually.

```{r}
library(SPARQLchunks)
endpoint <- "https://example.com/sparql"
query <- "
PREFIX ex: <http://example.com/schema/>
SELECT ?subject ?predicate ?object
WHERE {
  ?subject ?predicate ?object.
}
LIMIT 10
"

result_df <- sparql2df(endpoint, query)
```



